---
title: "Working with numbers"
author: "Karen C. Goncalves dos Santos"
callout-appearance: simple
date: "2 October 2023"
categories:
  - "R basics"
  - Numbers
  - Summarize
format: html
---

Functions for numeric data

## Sum, mean and median

You can obtain the sum, mean and median by calling the functions `sum()`, `mean()`, `median()`, respectively.
You need to include `na.rm=T` to obtain the estimate when you have missing values in the data.

```{r}
numeric_vector = c(iris$Sepal.Length, NA, NA)
sum(numeric_vector)
sum(numeric_vector, na.rm = T)
median(numeric_vector)
median(numeric_vector, na.rm = T)
mean(numeric_vector)
mean(numeric_vector, na.rm = T)
```

## Standard deviation, variance and interquantile range

For the data dispersion, you can use the functions `sd()` (standard deviation), `var()` (variance) and `IQR()` (interquantile range). Remember to use `na.rm=T`.

:::{.callout-note}
# Variance calculation

Variance = $\sum$ $( x - mean )^2$  / _N_

Where 

- _x_ is each and every value of the input vector
- _mean_ is the mean of the input vector
- _N_ is the length of the input vector 
:::

:::{.callout-note}

# Standard deviation calculation

Std.Dev = $\sqrt{Variance}$
:::

:::{.callout-note}
# Interquantile range calculation
`IQR(x) = quantile(x, 3/4) - quantile(x, 1/4)`

:::

```{r}
numeric_vector = c(iris$Sepal.Length, NA, NA)
sd(numeric_vector)
sd(numeric_vector, na.rm = T)
var(numeric_vector)
var(numeric_vector, na.rm = T)
#IQR(numeric_vector) Not accepted when there are NA values
IQR(numeric_vector, na.rm = T)
```


### Column-wise and row-wise summaries

You can calculate the sum and mean for each row/column of a numeric data frame with the functions:

- Sum: `colSums` and `rowSums`/`rowsum`
- Mean: `colMeans` and `rowMeans`
