---
title: "Overview data"
author: "Karen C. Goncalves dos Santos"
callout-appearance: simple
date: "1 November 2023"
format: html
categories:
  - "R basics"
  - Summarize
---

```{r, results='asis', echo = F}
source('styles.R')
```


## `str`, `summary` and other functions

### `str`

See data structure. Shows number of rows and columns in a data frame, classes of the columns, first few elements of each column.

```{r}
str(iris)
```

:::{.callout-note} 
You can add the option "max.level = X" if you know the object is nested (list inside of list inside of list) so that you don't print too many nested levels.
:::

### `head` or `tail`

Shows the beginning or the end of an object (you can specify the number of lines printed with `n = X`)

```{r}
head(iris, n = 7)
tail(iris, n = 3)
```

## `summary`

Print column-wise info about data frames, shows relevant data from tests, such as `aov` or t.test.

```{r}
summary(iris)
```

## Dimensions

- `dim(x)` 
	- Dimensions of a dataset: For a data frame it is c(1, 2), for a vector it is c(1).
-`length(x)` 
	- number of elements in a vector, or columns in a data frame or elements in a list
- `ncol` or `nrow` 
	- number of columns/rows in a data frame
- `names` 
	- Names of columns of a data frame
	- Names of objects in a named list or named vector

```{r}
dim(iris)
length(iris)
length(iris[,1])

ncol(iris)
nrow(iris)

names(iris)

my_numbers = 1:3; names(my_numbers) = c("A", "B", "C")
names(my_numbers)
```
