---
title: "Open files"
author: "Karen C. Goncalves dos Santos"
callout-appearance: simple
date: "3 November 2023"
categories:
  - "R basics"
  - Input files
format: html
---

```{r, results='asis', echo = F}
source('styles.R')
```

## Basic `read.XXXX` and tidyverse `read_XXXX`

### `read.XXX`

These functions are available without loading any package

```{r, echo=F, message=F, results='asis'}
library(tidyverse, quietly = T)

cheat_sheet = data.frame(
	Function = c('Function', 'read.csv', 'read.csv2', 'read.delim', 'read.delim2'),
	Column_separation = c('Column separation', '","', '";"', '"\\t" (tab)', '"\\t" (tab)'),
	Decimal_separation = c('Decimal separation', '"."', '","', '"."', '","')
)
caption = "Function names and their assumptions"
colNames = cheat_sheet[1, ]

cheat_sheet$Column_separation[2:nrow(cheat_sheet)] = 
	paste0("sep = ", cheat_sheet$Column_separation[2:nrow(cheat_sheet)])
cheat_sheet$Decimal_separation[2:nrow(cheat_sheet)] = 
	paste0("dec = ", cheat_sheet$Decimal_separation[2:nrow(cheat_sheet)])

cheat_sheet[-1,] %>%
	kableExtra::kable(
		format = "html", 
		col.names = colNames, row.names = F) %>% 
	kableExtra::kable_material_dark(lightable_options = "hover", html_font = "arial-narrow")

```

### `read_XXX`

These functions are available after loading `tidyverse`. They are the equivalent of the `read.XXXX`, but are faster and so better for big datasets.

Note that they produce `tibbles`, not `data.frames`. Though similar, `tibbles` do not have row names. If your first column has no name because it contains row names, R will replace the empty cell with `...1`.

You use `col_names` to indicate if there are column names or to give a vector with the column names.

```{r, echo=F, message=F, results='asis'}
library(tidyverse, quietly = T)

cheat_sheet = data.frame(
	Function = c('Function', 'read_csv', 'read_csv2', 'read_delim', 'read_delim2'),
	Column_separation = c('Column separation', '","', '";"', '"\\t" (tab)', '"\\t" (tab)'),
	Decimal_separation = c('Decimal separation', '"."', '","', '"."', '","')
)
caption = "Function names and their assumptions."
colNames = cheat_sheet[1, ]

cheat_sheet$Column_separation[2:nrow(cheat_sheet)] = 
	paste0("sep = ", cheat_sheet$Column_separation[2:nrow(cheat_sheet)])
cheat_sheet$Decimal_separation[2:nrow(cheat_sheet)] = 
	paste0("dec = ", cheat_sheet$Decimal_separation[2:nrow(cheat_sheet)])

cheat_sheet[-1,] %>%
	kableExtra::kable(
		format = "html", 
		col.names = colNames, row.names = F) %>% 
	kableExtra::kable_material_dark(lightable_options = "hover", html_font = "arial-narrow")

```