---
title: "Sequences and repetitions"
author: "Karen C. Goncalves dos Santos"
callout-appearance: simple
categories:
  - "R basics"
  - Numbers
date: "3 October 2023"
format: html
---

```{r, results='asis', echo = F}
source('styles.R')
```


## `seq()` and `rep` fucntions

### `seq`

Use the command `seq` to get a sequence of numbers. You can specify the interval between every number shown. The function is writen as:
`seq(FROM, TO, INTERVAL)`

```{r}
seq(1, 2, 0.1)

```

If you want to see every integer between two numbers you can use `:` or you can omit the interval in the `seq` command.

```{r}
# Get very integer between 0 and 10
seq(0, 10)
0:10

# Get every integer between 0 and 10 in decreasing order
seq(10, 0)
10:0
```

### `rep`

Syntax: `rep(X, TIMES, EACH)`

- X - what you want to repeat: vector of characters, numbers or logical (T/F)
- TIMES: repeat how many times
- EACH: if there is more than one element in WHAT, you repeat the first, then the second...

```{r}
# repeat X (1 in this case) 4 times
rep(1, 4)

# repeat X (vector c(1, 2)) 4 times
rep(c(1, 2), 4)

# repeat the elements of the input 4 times each
rep(c(1, 2), each = 4)
```
